\SW: Crowd-sourcing the Discovery of Gravitational Lenses

### 2015-04-08 doc/sw-system.tex
Brief description of sims, and discussion of strategy, added to section 2.2.2 and flagged in blue for comment. Closes #106

Gravitational lenses are unfamiliar objects to volunteers who are new
to the site. New volunteers need to learn what lenses look like as quickly as
possible, so that they can contribute informative classifications. They also
need to learn what lenses do not look like, in order to reduce the false
positive detection rate. There are three primary mechanisms in the \sw system
for teaching the volunteers what to look for. These are, in the order in which
they are encountered, an inline tutorial, instant feedback,
and a ``Spotter's Guide.'' As well as this, we provide ``About'', ``FAQ'' and ``Science'' pages explaining the physics of gravitational lensing. While we expect that the insight from this static material should help volunteers make sense of the features in the images, we focus on the more dynamic, activity-based training early, when engaging new volunteers to participate takes priority.

@@

The second image viewed after the initial tutorial image is already a survey
image, in order to get the volunteers engaged in the real task as quickly as
possible. Training continues beyond the first image tutorial through
``training subjects'' inserted randomly into the stream.  These training
subjects are either simulated lenses (known as ``sims''), or survey images
that were expert-classified and found not to contain any
lens candidates (these images are known as ``duds''). The tutorial explains
that the volunteers  will be shown such training images. They are also
informed that they will receive instant feedback about their performance after
classifying (blind) any of these training subjects. Indeed, after a volunteer
finishes marking a training subject and hits ``Finished marking,'' a pop-up
message is generated, containing either positive feedback for a successful
classification (for example, ``Well done! You spotted a simulated lens,'' as
in \Fref{fig:screenshot}) or negative feedback for an unsuccessful one (for
example, ``There is no gravitational lens in this field!'').

\todo{PJM}{Insert simulations paragraph here. General description of sim generation. Principles: use real images, draw from realistic populations, strike balance between easy to see (for education) and realistically difficult (to push sensitivity). \#106}.

>>>>

Gravitational lenses are largely unfamiliar objects to the general public. 
New volunteers need to learn what lenses look like as quickly as possible, so
that they can contribute informative classifications. They also need to learn
what lenses do not look like, in order to reduce the false positive detection
rate. There are three primary mechanisms in the \sw system for teaching the
volunteers what to look for. These are, in the order in which they are
encountered, an inline tutorial, instant feedback on ``training images''
inserted into the stream of images presented to them, and a ``Spotter's
Guide.'' As well as this, we provide ``About'', ``FAQ'' and ``Science'' pages
explaining the physics of gravitational lensing. While we expect that the
insight from this static material should help volunteers make sense of the
features in the images, we focus on the more dynamic, activity-based training
early, when engaging new volunteers to participate takes priority.

@@

The second image viewed after the initial tutorial image is already a survey
image, in order to get the volunteers engaged in the real task as quickly as
possible. However, training then continues, beyond the first image tutorial,
through ``training subjects'' inserted randomly into the stream.  These
training subjects are either simulated lenses (known as ``sims''), or survey
images that were expert-classified and found not to contain any lens
candidates (these images are known as ``duds''). The tutorial explains that
the volunteers  will be shown such training images. They are also informed
that they will receive instant feedback about their performance after
classifying (blind) any of these training subjects. Indeed, after a volunteer
finishes marking a training subject and hits ``Finished marking,'' a pop-up
message is generated, containing either positive feedback for a successful
classification (for example, ``Well done! You spotted a simulated lens,'' as
in \Fref{fig:screenshot}) or negative feedback for an unsuccessful one (for
example, ``There is no gravitational lens in this field!'').

\new{Because we include a training set of images in amongst the survey ``test''
images in this way, \sw can be viewed as a {\it supervised learning} system. 
Consequently, we should expect to find lens candidates that look like the sims
that we put in the training set, and so the design of this training set is
quite important. The training images must look realistic, in two ways. First,
apart from the simulated lenses they contain, they must look like all the
other survey images -- otherwise they could be identified as training images
without the volunteer learning anything about what lenses look like. For this
reason, we make the sims by {\it adding in} simulated lensed features to real
images drawn from the survey. It also means that each dataset presented for
inspection in the \sw interface needs to come with its own training set.}

\new{Second, the lensed features themselves must be realistic; if they did not
resemble real lenses, we could not expect to find real lenses in the test
images. The details of how we generated the sims for the CFHTLS project are
given in \PaperTwo, where we also carry out performance tests relative to real
lens candidates in the survey fields. Here, we merely note the following
aspects to place this paper's results in context.}

\new{We start by selecting massive galaxies and groups in the CFHTLS catalogs,
by color, brightness and photometric redshift. We then assign plausible mass
distributions to these ``potential lenses:'' a singular isothermal ellipsoid
model for each massive galaxy, and an additional NFW profile dark matter halo
for groups and clusters. The mass distributions are centred, elongated and
aligned with the measured optical properties of the galaxies present. We then
draw a source from a plausible background population, of either galaxies (for
the group or galaxy lenses) or quasars (for the galaxy-scale lenses). The
source redshift, luminosity and size distributions are all chosen to match
those observed, while the source abundance is artificially increased such that
each potential lens has a high probability of having a source occur at a
position that will lead to it being  highly magnified. Source galaxies are
given plausible surface brightness profiles and ellipticities. We then compute
the predicted multiple images of the source using the \gravlens ray-tracing
code \citep{Keeton2000}.}

\new{At this point, the distribution of lens and source properties are quite
realistic. At this point, we chose to make several cuts, to reject sims that
were likely to be difficult to spot. While this ensured the educational value
of the sims when giving feedback, it meant that we should only {\it expect} to
detect real lens candidates that also pass these cuts. Any real lenses
candidates we find with properties outside the ranges of the sims may in some
sense be considered bonuses, a product of the {\it imagination} of the crowd
in extrapolating their understanding of gravitational lenses into a regime
where they have not seen any before. In the CFHTLS project, the most important
selection cuts were to only keep sims with a)  second brightest image brighter
that $i = 23$ and b) magnification greater than 20 (\PaperTwo, Table 1).}
